Supporting Information

### 2018-02-23 SupportingInformation.Rmd
adding aphid and plankton north sea results for fitting

---
title: "Supporting Information"
author: "Shyamolina Ghosh and Daniel Reuman"
date: "February 12, 2018"
output: pdf_document
---

@@

# Descriptions of data
Each data contributor writes one to a few paragraphs about their data.

# Testing the ranking and bivariate copula display function

This is just for testing the `getcopula` function.

<!--Source in function and a package--->
```{r}
source("getcopula.R")
require(copula)
require(mvtnorm)
```

Plot data from a Clayton copula, which has lower tail dependence:
```{r}
copC<-claytonCopula(8,2)
dC<-rCopula(1000,copC)
vC<-getcopula(d=dC,rankon=F,ploton=T)
```

Plot data from a Joe copula, which has upper tail dependence:
```{r}
copJ<-joeCopula(8,2)
dJ<-rCopula(1000,copJ)
vJ<-getcopula(d=dJ,rankon=F,ploton=T)
```

Get data from a bivariate normal distribution and take ranks and
then plot:
```{r}
sigma<-matrix(c(1,.7,.7,1),2,2)
dN<-rmvnorm(1000,mean=c(0,0),sigma=sigma)
vN<-getcopula(d=dN,rankon=T,ploton=T)
```

# Testing the bivariate model selection and goodness-of-fit pipeline

This is just for testing the `OurBiCopSelect` function. We test that
it runs and gives reasonable results by simulating data and seeing if 
you recover what kind of copula was used.

<!--Source in the function and a package, changeable constants-->
```{r}
source("OurBiCopSelect.R")
require(copula)
nbsmall<-100
pthresh<-0.2
nblarge<-1000 #for bootstrapping, see below
numdat<-500 #amount of artificial data
seed<-101
```

If real data come from a Clayton copula, independence should
be rejected, Clayton should be the AIC best, it should pass the
goodness of fit test, and the normal copula should fail the
goodness of fit test:
<!--Test for real data from a clayton-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
set.seed(seed)

families<-c(1,3:6,13,14,16)

#data are really from a clayton
cop<-claytonCopula(5,2)
dat<-rCopula(numdat,cop)
allresC<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresC
```

Same test, but data now come from a Gumbel copula:
<!--Test for real data from a gumbel-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
set.seed(seed)

#data are really from a gumbel
cop<-gumbelCopula(6,2)
dat<-rCopula(numdat,cop)
allresG<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresG
```

Same test, but data now come from a Frank copula:
<!--Test for real data from a frank-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
set.seed(seed)

#data are really from a frank
cop<-frankCopula(5,2)
dat<-rCopula(numdat,cop)
allresF<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresF
```

Same test, but data now come from a Joe copula:
<!--Test for real data from a joe-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
set.seed(seed)

#data are really from a joe
cop<-joeCopula(6,2)
dat<-rCopula(numdat,cop)
allresJ<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresJ
```

The goodness of fit test is quite slow when testing the fit of a
normal copula, and the results are not great: often the normal copula
cannot be rejected even when it is not the true copula, even though
a comparison of A/BICs strongly supports the true copula over the normal.
So I think the best way forward is to use `gofnormal=FALSE`.

# Testing the pipeline for examining tail dependence in bivariate data

>>>>

---
title: "Supporting Information"
author: "Shyamolina Ghosh and Daniel Reuman"
date: "February 12, 2018"
output: pdf_document
---
updated on `r Sys.Date()`

@@

```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
families<-c(1,3:10,13,14,16:20)
source("getcopula.R")
source("OurBiCopSelect.R")
source("FittingCopula_selective_loc.R")
```

# Descriptions of data
<!--Each data contributor writes one to a few paragraphs about their data.-->

## Soil C and N data

## Aphid data
<!--Shya to add some facts, make sure to name the species-->
   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
   * We threw out all locations that had fewer than 30 years data 
   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data

## North Sea plankton data

   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
   * We threw out all locations that had fewer than 45 years data
   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data

# Question 1 analyses for each dataset

## Soil C and N data
<!--Shya to add model selection analysis of Terry data here. You just have to move it over from All_analysis.Rmd and edit the code. in a few ways:
1) make the code get displayed at each step
2) you have added some columns to the InfCritRes table so that will come out differently
3) you can also report your result for tail dependence based on model averaging
4) stop after that (e.g., don't put in the stuff about the outcome function, etc.-->
**About the dataset :**  
Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).

**Plot the raw data:**

```{r load_data_Loecke, echo=F}
#Load your data - should be a data frame with 2 columns. Can contain NAs.
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")]
plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
rect(0,0,1,1)
```

**Copula plot of the data:**  
```{r rankandplot_Loecke, echo=F}
#convert to a copula and plot
v<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
```

**Bivariate copula model selection and goodness of fit pipeline:**  
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v)}
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20)
BivMS_res_Loecke<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=families_Loecke,level=0.05,AICBIC="AIC",
                         numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                         gofnormal=F,status=T)
```

```{r}
BivMS_res_Loecke
BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw # -ve value indicates UT dep?
#or we should discard this result because none of the chosen model is a good fit for this dataset?
```

## Aphid data
<!--Shya to add model selection analysis here for aphid data, first the species for which you are using count data, then the one for ff data, just the model selection stuff, we'll do the other stuff with the tail dependence asymmetry statistics later-->
**Bivariate copula model selection and goodness of fit pipeline:** 
```{r vivj_matrix, echo=F}
# Select any two location pair [i,j] for a copula of species sp
# This function gives you a matrix with vi and vj as two columns
vivj_matrix<-function(d_allsp,sp,i,j){

  ds1<-d_allsp[[sp]][[i]]
  ds2<-d_allsp[[sp]][[j]]
  #----------------------------
  a1<-ds1$Year[1]
  a2<-ds2$Year[1]
  a3<-ds1$Year[dim(ds1)[1]]
  a4<-ds2$Year[dim(ds2)[1]]
  year_s<-max(a1,a2)
  year_e<-min(a3,a4)
  ind_s1<-which(ds1$Year==year_s)
  ind_s2<-which(ds2$Year==year_s)
  ind_e1<-which(ds1$Year==year_e)
  ind_e2<-which(ds2$Year==year_e)
  ds1<-ds1[ind_s1:ind_e1,]
  ds2<-ds2[ind_s2:ind_e2,]
  # Omitting the years and data containing NA in either d1 or d2 
  #from both d1 and d2
  if(anyNA(ds1$Dat)==T | anyNA(ds2$Dat)==T){
    ind_na1<-which(is.na(ds1$Dat))
    ind_na2<-c(ind_na1,which(is.na(ds2$Dat)))
    ind_na<-unique(ind_na2)

    d1Dat<-ds1$Dat[-ind_na]
    d2Dat<-ds2$Dat[-ind_na]
    Years<-ds1$Year[-ind_na]
    d1<-data.frame(Year=Years,Dat=d1Dat)
    d2<-data.frame(Year=Years,Dat=d2Dat)
  } else {
    d1<-ds1
    d2<-ds2
  }
  #get ranks modified now
  vi<-pobs(d1$Dat)
  vj<-pobs(d2$Dat)
  Years<-d1$Year
  #-------------------------
  #n_datapt<-length(vi)
  #--------------------
  #plot(vi,vj,type="p")
  #-------------------------
  mat<-as.matrix(cbind(vi,vj))
  return(mat)  
}
```

```{r d_allsp_data_aphid, echo=F}
d_allsp_data<-function(d0){

  name.aphid<-c("Apple-grass aphid","Bird cherry-oat aphid","Black bean aphid", "Blackberry-cereal aphid","Blackcurrant-sowthistle aphid",
                "Corn leaf aphid","Currant-lettuce aphid","Damson-hop aphid","Grain aphid","Green spruce aphid",
                "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach-potato aphid",
                "Potato aphid","Rose-grain aphid","Shallot aphid","Sycamore aphid","Willow-carrot aphid")
  longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
  lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)
  loc_name<-c("Ayr","Broom's Barn","Dundee","Edinburgh","Hereford","Newcastle","Preston","Rothamsted",
              "Starcross","Writtle","Wye")
  d0[d0==-999]<-NA
  Years<-1976:2010
  d_allsp<-list() #this is a list with 20 elements, all the species
  for (sp_c in 1:20)
  {
    d_thissp<-list() #a list with 11 locations for the current species
    for (st_c in 1:11)
    {
      d_thissp[[st_c]]<-data.frame(Year=Years,Dat=d0[st_c,((sp_c-1)*35+1):(sp_c*35)])
    }
    d_allsp[[sp_c]]<-d_thissp
  }
  names(d_allsp)<-name.aphid

  return(d_allsp=d_allsp)

}

# Aphid raw data for count
d0_count<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))

# Aphid raw data for first flight
d0_firstflight<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))

# Aphid raw data for flight duration
d0_flightduration<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))

```

## Aphid data, count

```{r aphid_count, echo=F, cache=T, cache.extra=list(seed,d0_count)}
set.seed(seed)
cat(paste("start-time: ",Sys.time(),"\n"))
RES_aphid_count<-RES_single_sp(d_allsp=d_allsp_data(d0_count),sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
cat(paste("stop-time: ",Sys.time(),"\n"))
resloc<-"Results/"
saveRDS(RES_aphid_count,paste(resloc,file="AphidCopulaFit_selecloc_count_species_10.RDS",sep=""))
```

```{r}
RES_aphid_count
#no of locations showing greater lower tail dependence
sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)

#no of locations showing greater upper tail dependence
sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)<0,na.rm=T)

```

## Aphid data, first flight

```{r aphid_firstflight, echo=F, cache=T, cache.extra=list(seed,d0_firstflight)}
set.seed(seed)
cat(paste("start-time: ",Sys.time(),"\n"))
RES_aphid_firstflight<-RES_single_sp(d_allsp=d_allsp_data(d0_firstflight),sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
cat(paste("stop-time: ",Sys.time(),"\n"))
resloc<-"Results/"
saveRDS(RES_aphid_firstflight,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
```

```{r}
RES_aphid_firstflight
#no of locations showing greater lower tail dependence
sum((RES_aphid_firstflight$LTdep_AICw-RES_aphid_firstflight$UTdep_AICw)>0,na.rm=T)

#no of locations showing greater upper tail dependence
sum((RES_aphid_firstflight$LTdep_AICw-RES_aphid_firstflight$UTdep_AICw)<0,na.rm=T)

```

## North sea plankton data
<!--Shya also add as above-->
```{r d_allsp_data_plankton_north_sea, echo=F}
d_allsp_data_plankton_north_sea<-function(d0){

  name.plankton<-c('Calanus_I-IV','Para-Pseudocalanus_spp','Temora_longicornis','Acartia_spp_(unidentified)','Centropages_typicus',
                   'Oithona_spp','Echinoderm_larvae','Calanus_finmarchicus','Calanus_helgolandicus','Metridia_lucens','Decapoda_larvae_(Total)',
                   'Euphausiacea_Total','Thalassiosira_spp','Rhizosolenia_styliformis','Ceratium_fusus','Ceratium_furca','Ceratium_tripos',
                   'Ceratium_macroceros','Rhizosolenia_alata_alata','Pseudocalanus_elongatus_Adult','Nitzschia_delicatissima','Nitzschia_seriata')

  longs<-read.csv("Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
  lats<-read.csv("Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
  longs<-longs[[1]]
  lats<-lats[[1]]
  d0[d0==-999]<-NA
  Years<-1958:2013

  d_allsp<-list() #this is a list with 20 elements, all the species
  for (sp_c in 1:22)
  {
    d_thissp<-list() #a list with 11 locations for the current species
    d1<-d0[((sp_c-1)*26+1):(sp_c*26),]
    for (st_c in 1:26)
    {
      d_thissp[[st_c]]<-data.frame(Year=Years,Dat=d1[st_c,])
    }
    d_allsp[[sp_c]]<-d_thissp
  }
  names(d_allsp)<-name.plankton

  return(d_allsp=d_allsp)

}

d0_plankton_north_sea<-as.matrix(read.csv("Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))

```

```{r plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,d0_plankton_north_sea)}
set.seed(seed)
cat(paste("start-time: ",Sys.time(),"\n"))
RES_plankton_north_sea<-RES_single_sp(d_allsp=d_allsp_data_plankton_north_sea(d0_plankton_north_sea),sp=16,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45)
cat(paste("stop-time: ",Sys.time(),"\n"))
resloc<-"Results/"
saveRDS(RES_plankton_north_sea,paste(resloc,file="Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS",sep=""))
```

```{r}
RES_plankton_north_sea
#no of locations showing greater lower tail dependence
sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)

#no of locations showing greater upper tail dependence
sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)<0,na.rm=T)

```

# Question 2 analyses for each dataset

# Question 3 analyses for each dataset
